Bridging the Gap Between Visual and Analytical Machine Learning Testing

### 2023-10-25 report/icst24.tex
squash

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{framed}
\newcommand{\highlight}[1]{\begin{framed}%
  \noindent\emph{#1}
\end{framed}}

\definecolor{backcolor}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
  basicstyle=\ttfamily\footnotesize,
  keywordstyle=\color{blue},
  stringstyle=\color{red},
  commentstyle=\color{green},
  breakatwhitespace=true,
  breaklines=true,
  keepspaces=true,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  numbers=left,
  numberstyle=\tiny\color{gray},
  backgroundcolor=\color{backcolor},
}
\lstset{style=mystyle}
\usepackage{graphicx}
\usepackage{caption}

@@

\begin{description}
\item[RQ1.] How frequently are analytical tests formulated from visualisations created to test ML systems?

@@

\item[RQ2.] What patterns are frequently observed in VA pairs used to test ML systems?

@@

\item[RQ3.] Do the assertions capture all the information obtained from the corresponding visualisation?

@@

@@

\begin{table*}
\centering
\caption{Frequent examples of post-condition checks observed in this study.}
  \begin{tabular}{m{0.2\textwidth} m{0.4\textwidth} m{0.3\textwidth}}
\hline
\emph{\textbf{Visualisation}}&
\emph{\textbf{Post-condition check}}&
\emph{\textbf{Rationale}}\\
\hline
\includegraphics[width=\linewidth]{post-cond-01.png}&
\begin{lstlisting}[language=Python]
assert f1.gca().has_data()
\end{lstlisting}&
The assertion checks if the output of the visualisation code cell contains an image.\\
\includegraphics[width=\linewidth]{post-cond-02.png}&
\begin{lstlisting}[language=Python]
assert_almost_equal(sigmoid(-50), 0, delta=1e-10)
assert_almost_equal(sigmoid(0), 0.5, delta=1e-10)
assert_almost_equal(sigmoid(50), 1, delta=1e-10)
\end{lstlisting}&
The assertions are spot-checking the sigmoid activation function.\\
\includegraphics[width=\linewidth]{post-cond-03.png}&
\begin{lstlisting}[language=Python]
# The axis should contain 5 bars
assert len(ax.containers[0]) == 5

# One of the x tick labels should be "blue"
tick_text = [tick.get_text() for tick in ax.get_xticklabels()]
assert "blue" in tick_text
\end{lstlisting}&
The asserts are validating the correctness of the visualisation itself.\\
\hline
\end{tabular}
\label{tab:post-cond}
\end{table*}

@@

We mined public repositories from Github to collect $54,070$ Jupyter notebooks that contain the keyword ``assert'' in them. We use the Github advanced search syntax~\footnote{https://docs.github.com/en/search-github/searching-on-github/searching-code} to isolate Jupyter Notebook. Additionally, we include the ``assert'' keyword in the search query with an \texttt{AND} operator. The final search query is as follows: \texttt{filename:"Jupyter Notebooks" AND "assert"}.

To further reduce the sample size, we exclude notebooks that do not use popular ML or Data Science (DS) libraries. We perform a string search to identify notebooks that import at least one of the libraries. The string patterns are derived from the python module name used in the import statements. The regex search query is as follows: \texttt{tensorflow|dask|torch|sklearn|scipy|numpy|pandas}.

@@

\begin{table}
  \centering
  \caption{Observations from random samples of notebooks with various
  cell arrangement.}
  \begin{tabular}{l p{0.2\textwidth} p{0.2\textwidth}}
    \hline
    \emph{\textbf{ID}}&
    \emph{\textbf{Cell arrangement}} &
    \emph{\textbf{Observations}}\\
    \hline
    1 &
    Assertion cell above visualisation cell &
    Least likely to be related to the visualisation below.\\
    \hline
    \multicolumn{3}{l}{\textbf{Assertion and visualisation in the same code cell}}\\
    \hline
    2.1 &
    Assertion above visualisation code &
    More likely to be a pre-condition check.\\
    2.2 &
    Assertion below visualisation code &
    More likely to be a post-condition check.\\
    \hline
    \multicolumn{3}{l}{\textbf{Assertion cell below visualisation cell}}\\
    \hline
    3.1 &
    Assertion cell more than 10 lines &
    Better signal-to-noise ratio compared to Arrangement 2.1. However, More likely to be post-condition check or unrelated to the visualisation above.\\
    3.2 &
    Assertion cell has max 10 lines &
    Most likely to be related to the visualisation above.\\
    \hline
  \end{tabular}
  \label{tab:cell-arrangement}
\end{table}

@@

\begin{table}
  \centering
  \caption{Inclusion-Exclusion criteria used to identify related visualisation-assertion pairs.}\label{tab:exclusion-criteria}
  \begin{tabular}{l p{0.2\textwidth} r}
    \hline
    \textbf{ID} &
    \textbf{Exclusion Criteria} &
    \textbf{Num. excluded NBs}\\
    \hline
    \textbf{EC1} &
    The visualisation and assertion are not related &
    560\\
    \textbf{EC2} &
    Notebook is not ML or DS related &
    419\\
    \textbf{EC3} &
    The assertion is a post-condition check &
    201\\
    \textbf{EC4} &
    Notebook not entirely written in English &
    164\\
    \textbf{EC5} &
    Notebook is a duplicate &
    129\\
    \textbf{EC6} &
    Notebook contains incorrect code &
    34\\
    \hline
    \multicolumn{2}{c}{\emph{\textbf{Total number of NBs excluded:}}} &
    1507\\
    \hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \centering
  \caption{Testing patterns observed in the visualisation-assertion pairs.}
  \begin{tabular}{l p{0.3\textwidth} c c}
    \hline
    \textbf{ID} &
    \textbf{Description} &
    \multicolumn{2}{c}{\textbf{Count}}\\
    & & Yes & No\\
    \hline
    \textbf{P1} &
    Does the assertion use a comparison operator (such as \emph{eq, lt, gt})? &
    202 & 67\\
    \textbf{P2} &
    Does the assertion use magic thresholds? &
    101 & 168\\
    \textbf{P3} &
    Does the assertion use an external testing library (such as \texttt{numpy.testing}? &
    42 & 227\\
    \hline
  \end{tabular}
  \label{tab:testing-patterns}
\end{table}

\begin{table}
  \centering
  \caption{Most frequently occurring assertions that use a comparison operator.}
  \begin{tabular}{m{0.9\linewidth}}
    \hline
    \begin{lstlisting}[language=Python]
assert accuracy_score(y_test, pred) > 0.95
    \end{lstlisting}\\
    \begin{lstlisting}[language=Python]
assert HISTORY['test_accuracy'][-1] > 0.92
    \end{lstlisting}\\
    \begin{lstlisting}[language=Python]
assert np.mean(mean_rw_history[-10:]) > 10.
print("That's good enough for tutorial.")
    \end{lstlisting}\\
    \begin{lstlisting}[language=Python]
assert np.mean(metrics['dev_bleu'][-10:], axis=0)[1] > 35, "We kind of need a higher bleu BLEU from you. Kind of right now."
    \end{lstlisting}\\
    \begin{lstlisting}
assert np.mean(dev_recall_history[-10:]) > 0.85, "Please train for at least 85% recall on test set. You may need to change vectorizer model for that."
    \end{lstlisting}\\
    \hline
  \end{tabular}
  \label{tab:compare-op-asserts}
\end{table}

@@

An overwhelming majority of assertions compare the value of a performance metric---typically the accuracy or recall---against a ``magic threshold'' defined by the notebook author. This covers 46\% of the assertions with a comparison operator. It is unclear however, how the specified thresholds were obtained since the authors do not provide additional explanation in the notebook. The threshold could have been derived based on domain expertise. Another alternative could be to define them during the Requirements Engineering phase, through discussions with all involved stakeholders\cite{CITEME}.\luis{Agathe Balayn might have something on these lines}

@@

Most notebooks examined in this study do no use any external testing libraries. 227 or 84\% of the VA pairs used the Python \texttt{assert} statement to define the test directly. This aligns with our observations that majority of the assertions compare the value of a performance metric against a user defined threshold, and do not require the use of a specialised testing library.

@@

The results from RQ2 provide emprical evidence to support the use of visualisations to test various properties of ML systems.  Visualisations are an efficient and effective medium to present complex information, that is easier to interpret by humans with the aid of visual cues. Visualisations used to test ML system properties however, demand manual verification effort. Information gained from visualisations and subsequent decisions made based on this information, must be manually verified in every iteration of the ML development cycle.

@@

Assertions or analytical tests extracted from ML visualisations can reduce this manual verification effort. Assertions encapsulate the knowledge gained from the visualisation with regards to the property of the ML system which is currently under test. Formal assertions document the observations made by the AI practitioner regarding the model or the data, at a specific point in time. These observations should not be violated by the future model development cycles unless the requirements of the ML system are changed. Assertions scale across organizational or personnel changes as future AI practitioners can refer to the assertions to understand how the visualisations were previously interpreted.

Results for RQ1 indicate that combining ML visualisations with analytical assertions is an emerging testing practice within the ML community. However in results from RQ2, we see several examples of assertions that fail to capture all the information gained from its visual counter-part. Our observations indicate a need to bridge the gap between qualitative assessment of ML systems using visualisations and deriving analytical assertions from the visualisations. We find the need for automated tools that make this translation process easier for AI practitioners.

@@

The empirical analysis presented in this study uses Jupyter notebooks mined from Github. To the best of our knowledge, \cite{pimentel2019large} and \cite{quaranta2021kgtorrent} are the only two prior studies which provide a public dataset of Jupyter Notebooks. We faced several technical challenges while trying to access the notebooks provided by these authors and thus decided to mine our own data from Github. We invite the scientific community to extend our catalogue of visual and analytical testing patterns, using the corpus of notebooks provided by the above authors.

In Section~\ref{sec:data-collect}, we intentionally keep the search criteria on Github broad by only checking for the appropriate filetype and the presence of the ``assert'' keyword. Besides being used to define tests in the Python programming language, ``assert'' also appears in several external python libraries that provide a testing interface (for instance nose, unittest and numpy). Our approach provides captures a rich variety of assertion statements while preventing the need to craft a custom search query based on an enhaustive list of keywords that appear in all Python testing libraries.

@@

This study mined 54K Jupyter Notebooks that contained an assertion. Further analysis identified 1.7K notebooks which contained an assertion directly below a visualisation. Of the 1.7K notebooks, 551 notebooks contained a visualisation and assertion pair which were semantically related to each other. A deeper analysis of 90 Jupyter notebooks revealved 26 unique ML testing tasks which utilised a combination of visualisations and assertions.

>>>>

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{framed}
\usepackage{tabularx}
\usepackage{booktabs}
\newcommand{\highlight}[1]{\begin{framed}%
  \noindent\emph{#1}
\end{framed}}

\definecolor{backcolor}{rgb}{0.95,0.95,0.92}
\colorlet{punct}{red!60!black}
\definecolor{delim}{RGB}{20,105,176}
\colorlet{numb}{magenta!60!black}

\lstdefinestyle{mystyle}{
  language=Python,
  basicstyle=\ttfamily\footnotesize,
  keywordstyle=\color{delim},
  stringstyle=\color{punct},
  commentstyle=\color{numb},
  breakatwhitespace=true,
  breaklines=true,
  keepspaces=true,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=2,
  stepnumber=1,
  numbersep=3pt,
  captionpos=b,
  numbers=left,
  numberstyle=\tiny\color{gray},
  backgroundcolor=\color{backcolor}
}
\lstset{style=mystyle}

\lstset{emph={%  
    assert%
    },emphstyle={\color{delim}}%
}%

\usepackage{graphicx}
\usepackage{caption}

@@

\begin{description}
  \item[RQ1.] \textbf{How frequently are analytical tests formulated from visualisations created to test ML systems?}

@@

\item[RQ2.] \textbf{What patterns are frequently observed in VA pairs used to test ML systems?}

@@

\item[RQ3.] \textbf{Do the assertions capture all the information obtained from the corresponding visualisation?}

@@

The replication package for this study is made public on Figshare\footnote{The replication package can be found at this URL: https://figshare.com/s/9347228011999cba29f5}.

@@

\begin{table*}
\centering
\caption{Frequent examples of post-condition checks observed in this study.}
  \begin{tabular}{@{}m{0.2\textwidth} m{0.42\textwidth} m{0.28\textwidth}@{}}
\toprule
\emph{\textbf{Visualisation}}&
\emph{\textbf{Post-condition check}}&
\emph{\textbf{Rationale}}\\
\midrule
\includegraphics[width=\linewidth]{post-cond-01.png}&
\begin{lstlisting}[language=Python,belowskip=0pt,aboveskip=0pt]
assert f1.gca().has_data()
\end{lstlisting}&
The assertion checks if the output of the visualisation code cell contains an image.\\
\includegraphics[width=\linewidth]{post-cond-02.png}&
\begin{lstlisting}[language=Python,belowskip=0pt,aboveskip=0pt]
assert_almost_equal(sigmoid(-50), 0, delta=1e-10)
assert_almost_equal(sigmoid(0), 0.5, delta=1e-10)
assert_almost_equal(sigmoid(50), 1, delta=1e-10)
\end{lstlisting}&
The assertions are spot-checking the sigmoid activation function.\\
\includegraphics[width=\linewidth]{post-cond-03.png}&
\begin{lstlisting}[language=Python,belowskip=0pt,aboveskip=0pt]
# The axis should contain 5 bars
assert len(ax.containers[0]) == 5

# One of the x tick labels should be "blue"
tick_text = [tick.get_text() for tick in ax.get_xticklabels()]
assert "blue" in tick_text
\end{lstlisting}&
The asserts are validating the correctness of the visualisation itself.\\
\bottomrule
\end{tabular}
\label{tab:post-cond}
\end{table*}

@@

We mined public repositories from Github to collect $54,070$ Jupyter notebooks that contain the keyword ``assert'' in them. We use the Github advanced search syntax~\footnote{https://docs.github.com/en/search-github/searching-on-github/searching-code} to isolate Jupyter Notebook. Additionally, we include the ``assert'' keyword in the search query with an \texttt{AND} operator. The final search query is as follows: \texttt{"image/png" AND "assert" language:"Jupyter Notebook"}.\footnote{Collected on June 22, 2023}

To further reduce the sample size, we exclude notebooks that do not use popular ML or Data Science (DS) libraries. We perform a string search to identify notebooks that import at least one of the libraries. The string patterns are derived from the python module name used in the import statements. The regex search query is as follows: \texttt{tensorflow|dask|torch|sklearn|scipy|numpy| pandas}.

@@

\begin{table}
  \centering
  \caption{Observations from random samples of notebooks with various
  cell arrangement.}
  \begin{tabular}{@{}l p{0.13\textwidth} p{0.27\textwidth}@{}}
    \toprule
    \emph{\textbf{ID}}&
    \emph{\textbf{Cell arrangement}} &
    \emph{\textbf{Observations}}\\
    \midrule
    1 &
    Assertion cell above visualisation cell &
    Least likely to be related to the visualisation below.\\
    \midrule
    \multicolumn{3}{l}{\textbf{Assertion and visualisation in the same code cell}}\\
    \midrule
    2.1 &
    Assertion above visualisation code &
    More likely to be a pre-condition check.\\
    2.2 &
    Assertion below visualisation code &
    More likely to be a post-condition check.\\
    \midrule
    \multicolumn{3}{l}{\textbf{Assertion cell below visualisation cell}}\\
    \midrule
    3.1 &
    Assertion cell more than 10 lines &
    Better signal-to-noise ratio compared to Arrangement 2.1. However, More likely to be post-condition check or unrelated to the visualisation above.\\
    3.2 &
    Assertion cell has max 10 lines &
    Most likely to be related to the visualisation above.\\
    \bottomrule
  \end{tabular}
  \label{tab:cell-arrangement}
\end{table}

@@

\begin{table}
  \centering
  \caption{Inclusion-Exclusion criteria used to identify related visualisation-assertion pairs.}\label{tab:exclusion-criteria}
  \begin{tabularx}{0.47\textwidth}{@{}l X r@{}}
    \toprule
    \textbf{ID} &
    \textbf{Exclusion Criteria} &
    \textbf{\# excluded NBs}\\
    \midrule
    \textbf{EC1} &
    The visualisation and assertion are not related &
    560\\
    \textbf{EC2} &
    Notebook is not ML or DS related &
    419\\
    \textbf{EC3} &
    The assertion is a post-condition check &
    201\\
    \textbf{EC4} &
    Notebook not entirely written in English &
    164\\
    \textbf{EC5} &
    Notebook is a duplicate &
    129\\
    \textbf{EC6} &
    Notebook contains incorrect code &
    34\\
    \midrule
    \multicolumn{2}{c}{\emph{\textbf{Total number of NBs excluded:}}} &
    1507\\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}
  \centering
  \caption{Testing patterns observed in the visualisation-assertion pairs.}
  \begin{tabularx}{0.45\textwidth}{@{}l X r r@{}}
    \toprule
    \textbf{ID} &
    \textbf{Description} &
    \multicolumn{2}{c}{\textbf{Count}}\\
    & & Yes & No\\
    \midrule
    \textbf{P1} &
    Does the assertion use a comparison operator (such as \emph{eq, lt, gt})? &
    202 & 67\\
    \textbf{P2} &
    Does the assertion use magic thresholds? &
    101 & 168\\
    \textbf{P3} &
    Does the assertion use an external testing library (such as \texttt{numpy.testing})? &
    42 & 227\\
    \bottomrule
  \end{tabularx}
  \label{tab:testing-patterns}
\end{table}

\begin{table}
  \centering
  \caption{Most frequently occurring assertions that use a comparison operator.}
  \begin{tabular}{@{}m{0.9\linewidth}@{}}
    \toprule
    \begin{lstlisting}[language=Python,belowskip=0pt,aboveskip=0pt]
assert accuracy_score(y_test, pred) > 0.95\end{lstlisting}\\
    \begin{lstlisting}[language=Python,belowskip=0pt,aboveskip=0pt]
assert HISTORY['test_accuracy'][-1] > 0.92\end{lstlisting}\\
    \begin{lstlisting}[language=Python,belowskip=0pt,aboveskip=0pt]
assert np.mean(mean_rw_history[-10:]) > 10.
print("That's good enough for tutorial.")\end{lstlisting}\\
    \begin{lstlisting}[language=Python,belowskip=0pt,aboveskip=0pt]
assert np.mean(metrics['dev_bleu'][-10:], axis=0)[1] > 35, "We kind of need a higher bleu BLEU from you. Kind of right now."\end{lstlisting}\\
    \begin{lstlisting}[language=python,belowskip=0pt,aboveskip=0pt]
assert np.mean(dev_recall_history[-10:]) > 0.85, "Please train for at least 85% recall on test set. You may need to change vectorizer model for that."\end{lstlisting}\\
    \bottomrule
  \end{tabular}
  \label{tab:compare-op-asserts}
\end{table}

@@

An overwhelming majority of assertions compare the value of a performance metric---typically the accuracy or recall---against a ``magic threshold'' defined by the notebook author\luis{we need to absolute values here (out of 269)}. This covers 46\% of the assertions with a comparison operator. It is unclear however, how the specified thresholds were obtained since the authors do not provide additional explanation in the notebook. The threshold could have been derived based on domain expertise. Another alternative could be to define them during the Requirements Engineering phase, through discussions with all involved stakeholders\cite{CITEME}.\luis{Agathe Balayn might have something on these lines}

@@

Most notebooks examined in this study do not use any external testing libraries. 227 or 84\% of the VA pairs use the low-level Python \texttt{assert} statement to define the test directly. This aligns with our observations that majority of the assertions compare the value of a performance metric against a user defined threshold, and do not require the use of a specialised testing library.

@@

The results from RQ2 provide empirical evidence to support the use of visualisations to test various properties of ML systems.  Visualisations are an efficient and effective medium to present complex information, that is easier to interpret by humans with the aid of visual cues. Visualisations used to test ML system properties however, demand manual verification effort. Information gained from visualisations and subsequent decisions made based on this information, must be manually verified in every iteration of the ML development cycle.

@@

Assertions or analytical tests extracted from ML visualisations can reduce this manual verification effort. Assertions encapsulate the knowledge gained from the visualisation with regards to the property of the ML system which is currently under test. Formal assertions document the observations made by the AI practitioner regarding the model or the data, at a specific point in time. These observations should not be violated by the future model development cycles unless the requirements of the ML system are changed. Assertions scale across organisational or personnel changes as future AI practitioners can refer to the assertions to understand how the visualisations were previously interpreted.

The results for RQ1 indicate that the combination of machine learning visualisations with analytical assertions is an emerging testing practice within the machine learning community. We found only 54,000 notebooks that contain the keyword ``assert'' and an image, while there are more than a million Jupyter notebooks available on GitHub. This suggests that the practice of assertion is still relatively uncommon.
We believe that this is due to a lack of knowledge among practitioners regarding testing approaches and the absence of adequate tooling to support the assertion of model properties. We argue that we need more guidelines on ML coding practices and that education and research of software engineering practices ought to be steered towards the specific challenges entailed from contributing to ML software.  We delve into this issue further in Section~\ref{sec:fw} \luis{I dont see this in fw and I wouldr remove this sentence because it breaks the flow of the narrative}.

\luis{Amother thought: there are still many visualisations without asserts. We believe in each visualisation there is an expectation that has not been formalised and that it will not be easily validated by other contributors. this is already in the motivation but we can frame it as a visible open problem here as well.}

Despite the efforts from some practitioners to automatically validate patterns from plots, results from RQ2 show that the classic software engineering testing practices are failing to adapt to the context of testing ML code. Very rarely assertions use dedicated API test methods. We argue that the root of this problem is two-fold: 1) the culture of testing ML code is still in its early stages and ML practitioners are not yet exploring software testing libraries, and 2) the existing testing APIs are not tuned to the specifics of ML code and practitioners have to fall back to basic built-in approaches.\luis{Side thought: we could also frame this is as: our results call for more tools that help testing ML code and that educational materials are needed to guide practitioners to testing best practices (I leave it to your judgement)}

\luis{some thought that is blended here but maybe we want to make it clear. ML testing is a task that requires practitioners to have a strong background in machine learning, software testing, and mathematics (or statistics). This reinforces the need to adjust the education of these professionals and to design tools that aim mitigate the lack of background in some of these directions.}

The wide adoption of ``magic'' thresholds (\luis{add numbers from results here}) is also concerning as it can hinder to the maintainability of ML projects. While in some cases there these thresholds come a results of domain expertise and experience acquired while working in a give project, we argue that such thresholds need to be properly documented. Our results show that there is a need to help practitioners keeping these decisions documented. This could be automated using a tool to flag undocumented decisions or that would suggest boilerplate documentation in Markdown or code annotations.

Additionally, in the results from RQ2\luis{this is RQ3}, we see several examples of assertions that fail to capture all the information gained from its visual counter-part. We argue that this can lead to a mismatch of expectations in future versions of the ML software: for example, another collaborator of the project might run the code with a new version of the dataset or a different learning technique and assume that if the assertions run successfully, the expected patterns from the visualisation are also being met. After delving into some examples, we also observe that deriving assertions from validations is far from trivial: basic visualisation patterns require complex mathematical techniques that might be out of the expertise of the ML developer. For example, the linear correlation observed with the scatter plot of Figure 6\luis{add ref} is clear to any non-expert but it had to be formalised (cf. Listing 11\luis{add ref}) using the metric coefficient of determination bounded to a sound threshold (0.6). Hence, we ought to bridge the gap between qualitative assessment of ML systems using visualisations and deriving analytical assertions from the visualisations. In other words, there is a need for automated tools that make this translation process easier for AI practitioners. In sum, our results showcase a gap in the current state of the art in software testing: it is failing to address the idiosyncrasies of ML code.

@@

The empirical analysis presented in this study uses Jupyter notebooks mined from Github. To the best of our knowledge, \cite{pimentel2019large} and \cite{quaranta2021kgtorrent} are the only two prior studies which provide a public dataset of Jupyter Notebooks. We faced several technical challenges while trying to access the notebooks provided by these authors and thus decided to mine our own data from Github. However, 24\% of the notebooks had to be excluded from the study since they were not related to ML or DS (Section~\ref{sec:identify-related-pairs} and Table~\ref{tab:exclusion-criteria}). We wish to extend our analysis to notebooks collected from Kaggle since Kaggle is predominantly targetted towards the ML and DS community.

In Section~\ref{sec:data-collect}, we intentionally keep the search criteria on Github broad by only checking for the appropriate filetype and the presence of the ``assert'' keyword. Besides being used to define tests in the Python programming language, ``assert'' also appears in several external python libraries that provide a testing interface (for instance nose\footnote{}, unittest and numpy). Our approach provides captures a rich variety of assertion statements while preventing the need to craft a custom search query based on an enhaustive list of keywords that appear in all Python testing libraries.

@@

In this study, we delved into the fusion of visualizations and analytical assertions in the testing of machine learning (ML) systems. Our investigation unveiled the emerging practice of crafting analytical assertions based on visual insights, a technique that bridges the qualitative and quantitative aspects of ML testing.

Testing ML systems in Jupyter Notebooks remains a burgeoning practice, with only a minority of notebooks adopting this approach. In total, we identified 269 pairs of visualizations and assertions that exhibited semantic correlation within 1.7k notebooks. Within these pairs, our analysis unveiled several key insights: nearly 40\% of the pairs included magic thresholds, 75\% of the assertions employed comparison operators, and 16\% of the assertions made use of external testing libraries. Furthermore, we observed that 22\% of the assertions did not utilize a comparison operator or a magic threshold.

Additionally, our study brought to light recurring testing patterns, underscoring the need yo enhance the alignment between visualizations and their corresponding assertions.
This research serves as a first step in comprehending the landscape of ML testing in Notebooks, emphasizing the potential for the development of more effective tools and practices to streamline the validation of visualizations and elevate the quality of analytical assertions.

\section{Future Work}\label{sec:fw}

This study underscores the imperative to enhance the current state of machine learning (ML) testing. Our observations have pointed to several potential future directions that could advance this field:

1. Tailored Assertion Tools: There is a clear need for assertion tools specifically designed to cater to the requirements of ML practitioners. These tools should provide assistance in identifying the most suitable visualizations for testing specific properties of ML systems. Customized solutions that streamline the process of connecting visual insights to analytical assertions would significantly benefit practitioners in this domain.

2. Automation of Assertions: The manual translation of insights derived from visualizations into analytical assertions presents a substantial challenge. Future work should focus on the development of automated tools that can assist AI practitioners in this translation process. By automating the generation of assertions from visual information, these tools can promote broader adoption of assertions in ML notebooks, reducing the burden of manual effort.

3. Dynamic Testing in Real-World Scenarios: Extending ML testing to dynamic real-world scenarios is a critical avenue for future exploration. This entails developing techniques and frameworks that allow assertions to adapt to evolving data distributions, changing model requirements, and continuous system updates. Research in this area will contribute to the development of ML systems that remain reliable and relevant throughout their operational lifecycles.
